---
title: "BATS_CO2"
author: "Raymond Infante"
date: "2023-03-23"
output: html_document
---

```{r setup, message=FALSE}
#Load Required packages
library(tidyverse)
library(seacarb)
library(performance)

#Import dataset
#(2 dataset to column names)
bats_bottle <- read_table("bats_bottle.txt", 
    col_names = FALSE, skip = 60)
bats_bottle_names <- read_csv("bats_bottle.txt", 
    skip = 59)


#Combine column names
colnames(bats_bottle) <- colnames(bats_bottle_names) 


#Check if the data and columns are correct
view(bats_bottle)
view(bats_bottle_names)

```
# Variables
yyyymmdd = Year Month Day   
decy   = Decimal Year     
time   = Time (hhmm)      
latN   = Latitude (Deg N) 
lonW   = Longitude (Deg W)
Depth  = Depth (m)                  
Temp   = Temperature ITS-90 (C)    
CTD_S  = CTD Salinity (PSS-78)      
Sal1   = Salinity-1 (PSS-78)        
Sig-th = Sigma-Theta (kg/m^3)       
O2(1)  = Oxygen-1 (umol/kg)          
OxFixT = Oxygen Fix Temp (C)        
Anom1  = Oxy Anomaly-1 (umol/kg)    
CO2    = dissolved inorganic carbon (umol/kg)              
Alk    = Alkalinity (uequiv)        
NO31   = Nitrate+Nitrite-1 (umol/kg)
NO21   = Nitrite-1 (umol/kg)        
PO41   = Phosphate-1 (umol/kg)      
Si1    = Silicate-1 (umol/kg)       
POC    = POC (ug/kg)                
PON    = PON (ug/kg)                
TOC    = TOC (umol/kg)                
TN     = TN (umol/kg)  
Bact   = Bacteria enumeration (cells*10^8/kg)   
POP    = POP (umol/kg)
TDP    = Total dissolved Phosphorus (nmol/kg)
SRP    = Low-level phosphorus (nmol/kg)
BSi    = Particulate biogenic silica (umol/kg)
LSi    = Particulate lithogenic silica  (umol/kg)
Pro    = Prochlorococcus (cells/ml)
Syn    = Synechococcus (cells/ml)
Piceu  = Picoeukaryotes (cells/ml)
Naneu  = Nanoeukaryotes (cells/ml)
/Quality flags
-999 = Missing or bad data
 0 = Less than detection limit



# Class questions
Can we quantify ocean acidification in the subtropical North Atlantic Ocean

1) Is surface ocean pco2 increasing? # we did this in class
2) Is surface ocean seawater ph decreasing?
3) Is surface ocean seawater saturation state with respect to arragonite decreasing? 


Question#1

```{r, echo=FALSE, message=FALSE, error=FALSE}
?carb
#carb(flag, var1, var2, S=35, T=25, Patm=1, P=0, Pt=0, Sit=0,
       # k1k2="x", kf="x", ks="d", pHscale="T", b="u74", gas="potential", 
        #warn="y", eos="eos80", long=1.e20, lat=1.e20)

#In the data we have TA, DIC (CO2), S, T, PT, sit ,but we dont have pressure

#First calculate pressure using TEOS-10

?gsw
# gsw_p_from_z(z,lat) (Pressures is going to be in dbar)

bats_co2 = bats_bottle %>% 
  mutate(P_gsw_dbar = gsw_p_from_z(Depth*-1, latN))
view(bats_co2)
#we now have all of the variables that we need to calculate the surface water chemistry at the BATS station, but that we need to be very careful about our units. 

# we now have In the data we have TA, DIC (CO2), S, T, PT, sit and pressure
#What are the units of these and waht does co2sys need?
#TA (uequiv) = umol/kg and we need mol/kg **** 
#=  Alk*10^-6
#DIC = umol/kg, we need mol/kg *****
#= DIC*10^-6
#S = is Practical Salinity and we will use EOS80 
# Sal1
#T = is in °C
# Temp
#Pt = umol/kg, we need mol/kg ******
# PO41*10^-6
#si1 = umol/kg, we need mol/kg *****
# Si1*10^-6
#P_dbar = dbar, we need bar ********
#P_dbar/10 or P_dbar*10^-1

#We will need to convert units scaling when using co2sys
#Flag 15 is alk and DIC

bats_co2sys = bats_co2 %>%
  filter(Alk!=-999, CO2!=-999, Sal1!=-999, Temp != -999, P_gsw_dbar!= -999, PO41!=-999, Si1!= -999) %>% 
  rename(DIC=CO2) %>% 
  mutate(carb(flag = 15, var1 = Alk*10^-6 , var2 = DIC*10^-6 , S=Sal1, 
     T=Temp, Patm=1, P=P_gsw_dbar*10^-1, 
     Pt=PO41*10^-6, Sit=Si1*10^-6, k1k2="l", 
     kf="pf", ks="d", pHscale="T"
     , b="u74", gas="potential", warn="y", 
     eos="eos80", long=360-lonW, lat=latN))
  view(bats_co2sys)
  
  
#flter for only hte surface ocean(only the upper 200m)
bats_co2sys_surf = bats_co2sys %>% 
  filter(Depth<100)
view(bats_co2sys_surface)

#Plot pco2 vs decy
bats_co2sys_surface %>% 
  ggplot(aes(x = decy, y = pCO2insitu))+
  geom_point(color = "#9d46bd", alpha = 0.3)+
  geom_line(alpha = 0.7)+
  theme_classic()+
  scale_x_continuous(limits = c(1990,2020))+
  labs(x = "Years", y = "pCO2", title = "pCO2 vs Years")+
  geom_smooth(method = "lm", color = "#ed6b5a")

bats_co2sys_surf %>% 
  filter(pCO2insitu<200) %>% 
  view()

pco2_model = lm(pCO2insitu~decy, data = bats_co2sys_surface)
summary(pco2_model)
#for every year it passes. pCO2 increase by 1.85 microatmosphere




#Checking our model:

pco2_model = lm(pCO2insitu~decy, data = bats_co2sys_surface)
summary(pco2_model)
anova(pco2_model)
#for every year it passes. pCO2 increase by 1.85 micro atmospshere

#Checking our models performance: Using the performance package.
check_model(pco2_model)
#model assumptions appear valid


#How do we plot our model predictions?
?predict
#This way we need to rename the columns
bats_co2sys_surf %>% 
  mutate(predict(pco2_model, interval = "prediction", level = 0.95))


#Base R method
bats_co2sys_surface_predict = cbind(bats_co2sys_surface, predict(pco2_model, interval = "confidence", level = 0.95))

bats_co2sys_surface_predict 

#Add prediction line to ggplot
bats_co2sys_surface_predict%>% 
  ggplot()+
  geom_point(mapping = aes(x = decy, y = pCO2insitu))+
  geom_line(aes(x = decy, y = fit), color = "purple")+
  geom_ribbon(aes(x = decy, ymin = lwr, ymax = upr), fill = "purple", alpha = 0.2)+
  theme_classic()
  
#There is a seasonal cycle in surface ocean pCO2 at BATS with higher pCO2 observed in late summer, to early fall and lower pco2 observed in late winter and early spring. There is a consistent. detectable (ANOVA, p<0.0001) annual increase in pco2 by 1.85+-0.07 uatm/year.
#Also show plot with model. The figure caption describes plot(points = data, line = model, shaded region = 95% confidence intervals.)

#For assignment:
#Put this graph cleaner.
#and do question 2 and 3 (pH and Aragonite). 
#for thursday
#try to include maps

####How can we improve our pCO2 predictions?
#What might be some valuable predictors of seawater pco2?
# Temperature - measure of atmospheric pco2, but also partial pressure of co2 in seawater
# Seasons -  months (winter vs summer)
# Years - annual increases in atmospheric CO2
# Dissolved Oxygen - photosynthesis and respiration impact both DO and CO2
# Salinity - impacts solubility
# Nutrients - also related through photosynthesis

#create a year and month column with BATS data
bats_co2sys_surf$year=
  as.numeric(substr(bats_co2sys_surface$yyyymmdd,1,4))
bats_co2sys_surf$month=
  as.numeric(substr(bats_co2sys_surf$yyyymmdd,5,6))

#we can build many models to fit our data
#more predictors = higher r^2
m1=lm(pCO2insitu~decy,data=bats_co2sys_surf)
m2=lm(pCO2insitu~year+month,data=bats_co2sys_surf)
m3=lm(pCO2insitu~Temp+year+month,data=bats_co2sys_surf)
summary(m1) #r^2=0.3
summary(m2) #r^2=0.4
summary(m3) #r^2=0.7
AIC(m1,m2,m3) #use AIC to select models (lower AIC = better model)

#for model development, I am interested in
#year, month, temp, sal, o2, no31,no21, po41
#that means I need to not have -999 in any of those

#bats_surf_sub=
#  bats_co2sys_surf %>% 
#  select(year,month,Temp,Sal1,`O2(1)`,`NO31`,
#         `NO21`,`PO41`) %>% 
#  replace_with_na_all(condition =~.x == -999)
#bats_surf_sub=bats_surf_sub[complete.cases(bats_surf_sub),]

#be cautious about removing missing data== see online lecture and notes about potential systematic biases
bats_surf_sub=
  bats_co2sys_surf %>%
  filter(year!=-999,month!=-999,Sal1!=-999,Temp!=-999,
         `O2(1)`!=-999,`NO31`!=-999,`NO21`!=-999,
         `PO41`!=-999)

#now we have all of our predictors without -999 missing data

step(lm(pCO2insitu~1,data=bats_surf_sub),
     direction="forward",
     trace = 1,
     scope = ~year+month+Sal1+Temp+`O2(1)`+`NO31`+`NO21`+ `PO41`)
  
model_AIC=lm(pCO2insitu ~ Temp + year + NO31 + Sal1 + NO21, data = bats_surf_sub)
summary(model_AIC)  
check_model(model_AIC)

bats_surf_sub_pred=
  cbind(bats_surf_sub, 
        predict(model_AIC, interval = 'confidence',level=0.95))

#add prediction line to ggplot
bats_surf_sub_pred %>% 
  ggplot()+
  geom_point(mapping=aes(x=decy,y=pCO2insitu))+
  geom_line(aes(decy,fit),color="purple")+
  geom_ribbon(aes(x=decy,ymin=lwr,ymax=upr),alpha=0.2,fill="purple")+
  theme_classic()

bats_surf_sub_pred %>% 
  ggplot()+
  geom_point(mapping=aes(x=decy,y=pCO2insitu-fit))+
  theme_classic()


#Compute correlation mix with pvalue
library(ggcorrplot)


#create correlation matrix
correlation_matrix = round(cor(bats_surf_sub),1)

#Correlation matrix p-value
corrp.mat = cor_pmat(bats_surf_sub)

#visualize our correlation matrix
ggcorrplot(correlation_matrix, method = "square")

#Principle components

bats_surf_sub2=
  bats_co2sys_surf %>%
  filter(year!=-999,month!=-999,Sal1!=-999,Temp!=-999,
         `O2(1)`!=-999,`NO31`!=-999,`NO21`!=-999,
         `PO41`!=-999) %>% 
  select(year, month, Sal1, Temp, `O2(1)` ,`NO31`,`NO21`,`PO41` )



ggcorrplot(correlation_matrix, method = "square")

pc2 = prcomp(bats_surf_sub2, center = T, scale. = T, na.rm =T)
pc
summary(pc)

library(ggfortify)
bats_surf_sub2$season = 
  ifelse(bats_surf_sub2$month ==12 | bats_surf_sub2$month ==1| bats_surf_sub2$month ==2|bats_surf_sub2$month ==3, "Winter","Not-winter")

view(bats_surf_sub2)
autoplot(pc2, loadings = T, loadings.label = T, color = 'season')

library(askgpt)
login()

askgpt("what is R")
```





