---
title: "BATS_CO2"
author: "Raymond Infante"
date: "2023-03-23"
output: html_document
---

```{r setup, message=FALSE}
#Load Required packages
library(tidyverse)
library(seacarb)

#Import dataset
#(2 dataset to column names)
bats_bottle <- read_table("bats_bottle.txt", 
    col_names = FALSE, skip = 60)
bats_bottle_names <- read_csv("bats_bottle.txt", 
    skip = 59)


#Combine column names
colnames(bats_bottle) <- colnames(bats_bottle_names) 


#Check if the data and columns are correct
view(bats_bottle)
view(bats_bottle_names)

```
# Variables
yyyymmdd = Year Month Day   
decy   = Decimal Year     
time   = Time (hhmm)      
latN   = Latitude (Deg N) 
lonW   = Longitude (Deg W)
Depth  = Depth (m)                  
Temp   = Temperature ITS-90 (C)    
CTD_S  = CTD Salinity (PSS-78)      
Sal1   = Salinity-1 (PSS-78)        
Sig-th = Sigma-Theta (kg/m^3)       
O2(1)  = Oxygen-1 (umol/kg)          
OxFixT = Oxygen Fix Temp (C)        
Anom1  = Oxy Anomaly-1 (umol/kg)    
CO2    = dissolved inorganic carbon (umol/kg)              
Alk    = Alkalinity (uequiv)        
NO31   = Nitrate+Nitrite-1 (umol/kg)
NO21   = Nitrite-1 (umol/kg)        
PO41   = Phosphate-1 (umol/kg)      
Si1    = Silicate-1 (umol/kg)       
POC    = POC (ug/kg)                
PON    = PON (ug/kg)                
TOC    = TOC (umol/kg)                
TN     = TN (umol/kg)  
Bact   = Bacteria enumeration (cells*10^8/kg)   
POP    = POP (umol/kg)
TDP    = Total dissolved Phosphorus (nmol/kg)
SRP    = Low-level phosphorus (nmol/kg)
BSi    = Particulate biogenic silica (umol/kg)
LSi    = Particulate lithogenic silica  (umol/kg)
Pro    = Prochlorococcus (cells/ml)
Syn    = Synechococcus (cells/ml)
Piceu  = Picoeukaryotes (cells/ml)
Naneu  = Nanoeukaryotes (cells/ml)
/Quality flags
-999 = Missing or bad data
 0 = Less than detection limit



# Class questions
Can we quantify ocean acidification in the subtropical North Atlantic Ocean

1) Is surface ocean pco2 increasing?
2) Is surface ocean seawater ph decreasing?
3) Is surface ocean seawater saturation state with respect to arragonite decreasing? 


```{r, echo=FALSE, message=FALSE, error=FALSE}
?carb
#carb(flag, var1, var2, S=35, T=25, Patm=1, P=0, Pt=0, Sit=0,
       # k1k2="x", kf="x", ks="d", pHscale="T", b="u74", gas="potential", 
        #warn="y", eos="eos80", long=1.e20, lat=1.e20)

#In the data we have TA, DIC (CO2), S, T, PT, sit ,but we dont have pressure

#First calculate pressure using TEOS-10

?gsw
# gsw_p_from_z(z,lat) (Pressures is going to be in dbar)

bats_co2 = bats_bottle %>% 
  mutate(P_gsw_dbar = gsw_p_from_z(Depth*-1, latN))
view(bats_co2)
#we now have all of the variables that we need to calculate the surface water chemistry at the BATS station, but that we need to be very careful about our units. 

# we now have In the data we have TA, DIC (CO2), S, T, PT, sit and pressure
#What are the units of these and waht does co2sys need?
#TA (uequiv) = umol/kg and we need mol/kg **** 
#=  Alk*10^-6
#DIC = umol/kg, we need mol/kg *****
#= DIC*10^-6
#S = is Practical Salinity and we will use EOS80 
# Sal1
#T = is in °C
# Temp
#Pt = umol/kg, we need mol/kg ******
# PO41*10^-6
#si1 = umol/kg, we need mol/kg *****
# Si1*10^-6
#P_dbar = dbar, we need bar ********
#P_dbar/10 or P_dbar*10^-1

#We will need to convert units scaling when using co2sys
#Flag 15 is alk and DIC

bats_co2sys = bats_co2 %>%
  filter(Alk!=-999, CO2!=-999, Sal1!=-999, Temp != -999, P_gsw_dbar!= -999, PO41!=-999, Si1!= -999) %>% 
  rename(DIC=CO2) %>% 
  mutate(carb(flag = 15, var1 = Alk*10^-6 , var2 = DIC*10^-6 , S=Sal1, 
     T=Temp, Patm=1, P=P_gsw_dbar*10^-1, 
     Pt=PO41*10^-6, Sit=Si1*10^-6, k1k2="l", 
     kf="pf", ks="d", pHscale="T"
     , b="u74", gas="potential", warn="y", 
     eos="eos80", long=360-lonW, lat=latN))
  view(bats_co2sys)
  
  
#flter for only hte surface ocean(only the upper 200m)
bats_co2sys_surface = bats_co2sys %>% 
  filter(Depth<100)
view(bats_co2sys_surface)


bats_co2sys_surface %>% 
  ggplot(aes(x = decy, y = pCO2insitu))+
  geom_point()+
  geom_line()

bats_co2sys_surface %>% 
  filter(pCO2insitu<200) %>% 
  view()

pco2_model = lm(pCO2insitu~decy, data = bats_co2sys_surface)
summary(pco2_model)

#For Homework:
#Make decy vs pco2 plot pretty
#Could include methods text for co2sys in lab report

#How do we check our model performance? 
#How do we plot our model predictions?

```

For the markdown do:
Intro - methods - results - discussion mini-lab report